### 2020-11-22 paper.tex
First commit.

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

%% Please use the following documentclass and journal abbreviations for preprints and final revised papers.

%% 2-column papers and preprints
\documentclass[gchron, manuscript]{copernicus}

\usepackage{upquote}

%% Journal abbreviations (please use the same for preprints and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Bone and Joint Infection (jbji)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

\begin{document}

\title{On the treatment of discordant detrital zircon U--Pb data}

% \Author[affil]{given_name}{surname}

\Author[1]{Pieter}{Vermeesch}

\affil[1]{Department of Earth Sciences, University College London, Gower Street, London WC1E 6BT}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Aman}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".

%% If authors contributed equally, please mark the respective author names with an asterisk, e.g. "\Author[2,*]{Anton}{Aman}" and "\Author[3,*]{Bradley}{Bman}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".

\correspondence{Pieter Vermeesch (p.vermeesch@ucl.ac.uk)}

\runningtitle{Treatment of discordant U--Pb data}

\runningauthor{Pieter Vermeesch}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
  Detrital zircon U--Pb geochronology is a staple of sedimentary
  provenance analysis and crustal evolution studies. Constructing
  detrital age spectra is straightforward for concordant
  \textsuperscript{206}Pb/\textsuperscript{238}U- and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb-compositions.  But
  unfortunately, many detrital U--Pb datasets contain a significant
  proportion of discordant analyses. This situation is expected to
  worsen in the future as a result of increasing analytical precision,
  which reveals ever smaller deviations from isotopic concordance.
  The analysis of discordant U--Pb data involves two largely arbitrary
  decisions.

  First, the analyst must choose whether to use the
  \textsuperscript{206}Pb/\textsuperscript{238}U- or the
  \textsuperscript{207}Pb/\textsuperscript{206}Pb-date. The
  \textsuperscript{206}Pb/\textsuperscript{238}U-method is more
  precise for young samples, whereas the
  \textsuperscript{207}Pb/\textsuperscript{206}Pb-method is better
  suited for old samples. However there is no agreement which `cutoff'
  should be used to switch between the two. This subjective decision
  can be avoided by using single grain concordia ages. These represent
  a kind of weighted mean between the
  \textsuperscript{206}Pb/\textsuperscript{238}U- and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb-methods, which
  offers better precision than either of the latter two methods.

  A second subjective decision is how to define the discordance cutoff
  between `good' and `bad' data. Discordance is usually defined as (1)
  the relative age difference between the
  \textsuperscript{206}Pb/\textsuperscript{238}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb dates.  However,
  this paper shows that several other definitions are possible as
  well, including (2) the absolute age difference; (3) the common-Pb
  fraction according to the Stacey-Kramers mantle evolution model; (4)
  the p-value of concordance; (5) the perpendicular logratio (or
  `Aitchison') distance to the concordia line; and (6) the logratio
  distance to the maximum likelihood composition on the concordia
  line.

  Applying these six discordance filters to a 10,000-grain dataset of
  detrital zircon U--Pb compositions reveals that: (i) the relative
  age discordance filter tends to suppress the young ($<1.5$~Ga) age
  components in U--Pb age spectra, whilst inflating the older
  ($>1.5$~Ga) age components; (ii) the absolute age difference and
  Stacey-Kramers discordance filters are more likely to reject old
  grains and less likely to reject young ones; (iii) the p-value based
  discordance filter has the undesirable effect of biasing the results
  towards the least precise measurements; (iv) the logratio-based
  discordance filters are most strict for Proterozoic grains, and more
  lenient for Phanerozoic and Archaean age components; (v) of all the
  methods, the logratio distance to the concordia composition produces
  the best results, in the sense that it yields nearly identical
  \textsuperscript{206}Pb/\textsuperscript{238}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb age spectra.  All
  the methods presented in this paper have been implemented in the
  \texttt{IsoplotR} toolbox for geochronology.
\end{abstract}

%\copyrightstatement{TEXT} %% This section is optional and can be used for copyright transfers.

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}

Detrital zircon U--Pb geochronology is a widely used tool for
sedimentary provenance analysis, crustal evolution studies and maximum
depositional age estimation.  In all these applications, inferences
are based on the frequency distribution of U--Pb age \emph{estimates},
which are derived from isotopic ratio \emph{measurements}. Several
mathematical and statistical operations are required to extract an age
spectrum from a table of U--Pb isotope ratio measurements. There
exists a lack of consensus among the detrital zircon geochronology
community about some of these steps.  Two outstanding questions are:

\begin{enumerate}
\item Which age estimate to use? It is widely recognised that
  \textsuperscript{206}Pb/\textsuperscript{238}U age estimates offer
  the optimal accuracy and precision at the young end of the age
  spectrum, whereas the
  \textsuperscript{207}Pb/\textsuperscript{206}Pb method is better
  suited for older samples. However the cutoff between the two clocks
  varies between studies, with values ranging from 800~Ma to 1.5~Ga
  \citep{gehrels2011,spencer2016}.
\item How to treat discordant data? The most reliable age constraints
  are obtained from zircons whose
  \textsuperscript{206}Pb/\textsuperscript{238}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb ages are
  statistically indistinguishable from each other. U--Pb compositions
  that fulfil this requirements are `concordant'. Those that violate
  it are `discordant'. It is not obvious how to quantify the
  discordance of the \textsuperscript{206}Pb/\textsuperscript{238}U
  and \textsuperscript{207}Pb/\textsuperscript{206}Pb systems. Most
  studies define discordance as the relative age difference between
  the \textsuperscript{206}Pb/\textsuperscript{238}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb ages, but some
  advocate the use of statistical hypothesis tests and p-values to
  quantify discordance \citep{spencer2016}. And even when a
  discordance definition has been agreed upon, there are many ways to
  choose the discordance cutoff. For example, the relative age
  discordance threshold may vary between 10\% and 30\%
  \citep{gehrels2011}.
\end{enumerate}

This paper addresses both of these issues. Section~\ref{sec:whichage}
shows that single-grain concordia ages \citep{ludwig1998} can be used
to avoid the arbitrary cutoff between the
\textsuperscript{206}Pb/\textsuperscript{238}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb
methods. Section~\ref{sec:discordance1} compares and contrasts
existing discordance filters based on age disparity and p-values. It
shows that the p-value definition hurts both the accuracy and
precision of detrital geochronology. Section~\ref{sec:discordance2}
proposes three new ways to quantify discordance, based directly on
U--Pb compositions rather than the ages calculated therefrom. Two of
these new definitions are based on a logratio distance that is widely
used in compositional data analysis \citep{aitchison1986}. Finally,
Section~\ref{sec:application} applies the six discordance filters to a
compilation of detrital zircon U--Pb data. Although the true age
distribution of this dataset is unknowable, the results suggest that
the logratio based discordance filters produce the most accurate
results.

\section{Which age to choose?}\label{sec:whichage}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{TW1500.pdf}
  \caption{Illustrative Tera-Wasserburg concordia diagram with a
      concordant and discordant measurement. $t_{68}$ marks the
      \textsuperscript{206}Pb/\textsuperscript{238}U age, $t_{76}$ the
      \textsuperscript{207}Pb/\textsuperscript{206}Pb age, and $t_{c}$
      the concordia age. Measurement~1 is concordant because its
      estimates for $t_{68}$, $t_{76}$ and $t_{c}$ are
      identical. Measurement~2 is discordant because the three
      estimates disagree. The concordia age is the most likely age
      given the analytical uncertainties. It falls between the other
      two age estimates, and offers the best analytical precision of
      the three.
  }
  \label{fig:concordia}
\end{figure}

The U--Pb method is based on three separate chronometers:
\textsuperscript{206}Pb/\textsuperscript{238}U,
\textsuperscript{207}Pb/\textsuperscript{235}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb. The half-life of
\textsuperscript{235}U is more than six times shorter than that of
\textsuperscript{238}U, and \textsuperscript{235}U is more than 100
times less abundant than \textsuperscript{238}U. For these two
reasons, little \textsuperscript{207}Pb has been produced during the
last billion years of Earth history compared to
\textsuperscript{206}Pb. Consequently, the
\textsuperscript{207}Pb/\textsuperscript{235}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb methods are less
precise than the \textsuperscript{206}Pb/\textsuperscript{238}U method
during the Phanerozoic and Neoproterozoic.

However, during earlier stages of Earth's history,
\textsuperscript{235}U was significantly more abundant than it is
today. The \textsuperscript{238}U/\textsuperscript{235}U ratio was
$\sim$60 at 1Ga, $\sim$26 at 2Ga, $\sim$11 at 3Ga, and $\sim$5 at 4Ga.
Due to the greater abundance of \textsuperscript{235}U in this past,
and because it decays much faster than \textsuperscript{238}U, the
precision of the \textsuperscript{207}Pb/\textsuperscript{235}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb clocks exceeds that of
the \textsuperscript{206}Pb/\textsuperscript{238}U method during the
Palaeoproterozoic and Archaean. The gradual shift in sensitivity
between the two chronometers is visible in the slope of a
Tera-Wasserburg concordia line, which is steep at old ages (high
\textsuperscript{207}Pb/\textsuperscript{206}Pb gradient w.r.t. time)
and shallow at young ages (low
\textsuperscript{238}U/\textsuperscript{206}Pb gradient
w.r.t. time).

Most published detrital zircon U--Pb studies switch from
\textsuperscript{206}Pb/\textsuperscript{238}U to
\textsuperscript{207}Pb/\textsuperscript{206}Pb at some point during
the Proterozoic. Unfortunately there are two problems with such a
switch. First, it requires the selection of a discrete discordance
cutoff between the two methods. If this cutoff differs between two
studies (which it often does), then this complicates the
intercomparison of their respective age spectra. Second, the sudden
switch between the \textsuperscript{206}Pb/\textsuperscript{238}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb clocks is often marked
by a discrete step in the age spectrum \citep{puetz2018}. This step is
entirely artificial and obscures any geologically significant events
that might occur around the same time.

Both of these problems can be solved by using `hybrid' concordia ages
instead of `pure' \textsuperscript{206}Pb/\textsuperscript{238}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb ages. Concordia ages
are defined by \citet{ludwig1998} as the `most likely' (in a
statistical sense) U--Pb age given the isotopic ratio composition and
its analytical uncertainty (Figure~\ref{fig:concordia}).  Let $r_{75}$
and $r_{68}$ be the measured
\textsuperscript{207}Pb/\textsuperscript{235}U and
\textsuperscript{206}Pb/\textsuperscript{238}U ratios, respectively,
and let $\sigma[r_{68}]^2$, $\sigma[r_{76}]^2$,
$\sigma[r_{68},r_{76}]$ be their (co)variances.  Then the concordia
age $t_c$ is obtained by numerically minimising the sum of squares
$S$:
\begin{equation}
  S = \left[
    \begin{array}{@{}c@{}}
      r_{75} - \exp(\lambda_{235}t_c) + 1 \\
      r_{68} - \exp(\lambda_{238}t_c) + 1
    \end{array}
    \right]^T\left[
    \begin{array}{@{}cc@{}}
      \sigma[r_{75}]^2 & \sigma[r_{75},r_{68}] \\
      \sigma[r_{75},r_{68}] & \sigma[r_{68}]^2
    \end{array}
    \right]^{-1}
  \left[
    \begin{array}{@{}c@{}}
      r_{75} - \exp(\lambda_{235}t_c) + 1 \\
      r_{68} - \exp(\lambda_{238}t_c) + 1      
    \end{array}
    \right]
  \label{eq:S}
\end{equation}

The single grain concordia age combines the chronometric power of the
\textsuperscript{206}Pb/\textsuperscript{238}U and
\textsuperscript{207}Pb/\textsuperscript{206}Pb systems. For young
($<$1~Ga) samples, the concordia age is nearly identical to the
\textsuperscript{206}Pb/\textsuperscript{238}U age. For old samples
($>$2~Ga) it approaches the
\textsuperscript{207}Pb/\textsuperscript{206}Pb age. For Proterozoic
samples, the concordia age gradually shifts form the
\textsuperscript{206}Pb/\textsuperscript{238}U to the
\textsuperscript{207}Pb/\textsuperscript{206}Pb age. Using concordia
ages removes the need for an arbitrary cutoff between the two
chronometers. An additional advantage is that the concordia age offers
better precision than the
\textsuperscript{206}Pb/\textsuperscript{238}U and the
\textsuperscript{207}Pb/\textsuperscript{206}Pb chronometer (or the
\textsuperscript{207}Pb/\textsuperscript{235}U for that matter)
(Figure~\ref{fig:precision}).

\begin{figure}[t]
  \includegraphics[width=8.3cm]{agerr.pdf}
  \caption{Predicted uncertainties of the
    \textsuperscript{207}Pb/\textsuperscript{235}U,
    \textsuperscript{206}Pb/\textsuperscript{238}U,
    \textsuperscript{207}Pb/\textsuperscript{206}Pb and concordia
    ages for a synthetic dataset with a constant uranium
    concentration. Dwell times and detector sensitivities were
    chosen so as to yield results that are similar to those obtained
    from real data. The concordia age (solid line) always offers the
    best precision. See the Appendix for further details about the
    calculations behind this figure.}
  \label{fig:precision}
\end{figure}

\section{Discordance filters: old definitions}\label{sec:discordance1}

Discordance may be caused by the addition of common Pb, the partial
removal of Pb during high grade metamorphism, or the mixing of
diachronous growth zones during micro-analysis. Most commonly,
discordance is defined in terms of the relative difference between the
\textsuperscript{206}Pb/\textsuperscript{238}U age and the
\textsuperscript{207}Pb/\textsuperscript{206}Pb age
\citep{gehrels2011}:
\begin{equation}
  d_r = 1 - t_{68}/t_{76}
  \label{eq:dr}
\end{equation}

However other definitions are possible as well. For example, one could
also define discordance in terms of absolute age differences
\citep{puetz2018}:
\begin{equation}
  d_t = t_{76} - t_{68}
  \label{eq:dt}
\end{equation}

A third option is to define discordance in terms of U--Pb compositions
rather than ages. \citet{spencer2016} advocate using p-values to
assess concordance. In the context of single grain concordia ages, the
p-value is the probability that the sum of squares $S$
(Equation~\ref{eq:S}) exceeds the observed value under a chi-square
distribution with two degrees of freedom:
\begin{equation}
  d_p = \mbox{Prob}\left(s > S | S \sim \chi^2_2
    \right)
  \label{eq:dp}
\end{equation}

Detrital zircon U--Pb data can be filtered by removing all
measurements whose discordance values exceed a certain threshold
value. Typical cutoff values for $d_r$ are 10--30\%
\citep{gehrels2011}, whereas $d_p$ is generally set to 5\%
\citep{spencer2016}. Different discordance criteria produce different
U--Pb age spectra. For example, a relative age cutoff will
preferentially remove young grains whereas an absolute age cutoff is
more likely to remove old grains (Figure~\ref{fig:agediscordance}).

\begin{figure}[t]
  \includegraphics[width=8.3cm]{TW-option-12.pdf}
  \caption{Discordance filters using a 200~Ma absolute age cutoff
    (dashed lines) and a 20\% relative age cutoff (dotted
    lines). The Tera-Wasserburg concordia diagram is plotted in
    logarithmic space to provide a more balanced view of the old and
    young ends of the time scale.
  }
  \label{fig:agediscordance}
\end{figure}

The p-value definition affects grains differently depending on their
analytical precision \citep{nemchin2005}. For example, consider a
1.5~Ga zircon that is 1\% discordant. If this grain were analysed by
LA-ICP-MS with an analytical precision of 2\%, say, then it would pass
the chi-square test and be accepted as being concordant. However, if
that same grain were analysed by TIMS with a precision of 0.2\%, then
the p-value criterion would reject it as being discordant. It seems
fundamentally unfair that an imprecise analytical method would be
favoured over a precise one (Figure~\ref{fig:TIMSvsLAICPMS}). This is
a pertinent problem because technical innovations are increasing the
precision of all analytical approaches to U--Pb geochronology.  As
precision improves, so does the ability to detect even small degrees
of discordance. Using the p-value criterion, there may come a time
when no detrital zircon passes this filter.

A final argument against the p-value discordance criterion is that it
biases against old U--Pb ages. This is because old zircon contains
more radiogenic Pb than young zircon does. Therefore the analytical
precision of the isotopic ratio measurements tends to be better for
old grains than it is for young ones. Consequently, the chi-square
test has greater power \citep[\emph{sensu}][]{cohen1992} to reject
them. In conclusion, p-value based discordance filters are
fundamentally flawed. Despite their appeal as `objective' tools for
statistical decision making, formalised hypothesis tests such as
chi-square are rarely useful in geology. For the same reason, the
widely used MSWD \citep[Mean Square of the Weighted
  Deviates,][]{mcintyre1966} statistic (which is just $S$/2 in this
case) should be used with caution. This is because, like p-values,
also MSWD cutoffs punish precise datasets in favour of imprecise
ones. Note that this caveat also goes against the recommendations of
\citet{spencer2016}.

\begin{figure}
  \includegraphics[width=8.3cm]{TIMSvsICPMS.pdf}
  \caption{Application of the flawed p-value discordance criterion
    to two synthetic measurements by TIMS (left) and LA-ICP-MS
    (right).  The precise TIMS measurement is labelled as discordant
    even though it plots closer to the concordia line than the
    imprecise LA-ICP-MS measurement, which is labelled as
    concordant.
  }
  \label{fig:TIMSvsLAICPMS}
\end{figure}

\section{Discordance filters: new definitions}\label{sec:discordance2}

\begin{figure}
  \includegraphics[width=8.3cm]{TW-option-linear-3.pdf}
    \caption{Using the \citet{stacey1975} common Pb model as a
      discordance criterion. This criterion assumes that the
      discordance is caused by linear mixing (hence, the linear scale
      of this Tera-Wasserburg plot) between radiogenic Pb
      (intersections of the mixing lines with concordia) and common Pb
      (intersection of the mixing lines with the vertical axis). The
      dashed line marks the 20\% ($= d_1/[d_1+d_2]$) discordance
      cutoff. This discordance filter, which must be applied
      \emph{before} making any actual common Pb correction, is more
      forgiving for young grains than it is for old grains. In this
      respect, it has the opposite effect of the relative age filter
      shown in Figure~\ref{fig:agediscordance}.
    }
    \label{fig:SK}
\end{figure}

Section~\ref{sec:discordance1} reviewed three existing discordance
definitions.  This Section will introduce three new ones.  None of the
definitions discussed thus far encode any information about the
geological mechanisms behind the discordance. As explained in
Section~\ref{sec:intro}, common Pb is one of the most likely causes of
discordance. Using a mantle evolution model \citep[e.g.][]{stacey1975}
to approximate the isotopic composition of this common Pb, discordance
can be defined as:
\begin{equation}
  d_{sk} = 1 - \left[\frac{{}^{238}U}{{}^{206}Pb}\right]\bigg/\left[\frac{{}^{238}U}{{}^{206}Pb}\right]^\ast
  \label{eq:dsk}
\end{equation}

\noindent where $[{}^{238}U/{}^{206}Pb]^\ast$ is the
\textsuperscript{238}U/\textsuperscript{206}Pb-ratio of the
intersection between concordia and a straight line connecting the
\textsuperscript{238}U/\textsuperscript{206}Pb--\textsuperscript{207}Pb/\textsuperscript{206}Pb
measurement to the inferred mantle composition (Figure~\ref{fig:SK}).

The common Pb definition of discordance is more forgiving for young
grains than it is for old ones. Importantly, if the discordance is
caused by common Pb, then the
\textsuperscript{206}Pb/\textsuperscript{238}U,
\textsuperscript{207}Pb/\textsuperscript{206}Pb and concordia age
estimates are all negatively biased with respect to the true
age. However this bias can be removed by applying a common-Pb
correction \emph{after} the data have been filtered.

Each discordia definition that we have studied thus far is expressed
in different units. For the absolute age definition, degrees of
discordance are expressed in units of time (ranging from 0 to
4.5~Ga). The relative age definition uses fractions of time (ranging
from $-\infty$ to 1). The p-value definition expresses discordance in
terms of probability (ranging from 0 to 1). And the \citet{stacey1975}
definition uses fractions of ratios (ranging from $-\infty$ to
1). None of these scales is particularly intuitive or natural. They
certainly do not match the usual definition of \emph{distance} in the
geographical sense of the word.

To address this issue, it is useful to subject the U--Pb isotopic
ratio data to a logarithmic transformation. So instead of analysing
our data on a conventional Tera-Wasserburg concordia diagram, all
calculations can be done in
log(\textsuperscript{207}Pb/\textsuperscript{206}Pb) vs.
log(\textsuperscript{238}U/\textsuperscript{206}Pb) space. The
advantage of this transformation is that it produces values that are
free to range from $-\infty$ to $+\infty$. Within this infinite
dataspace, the Euclidean distance metric can be safely applied.

\begin{figure}
  \includegraphics[width=8.3cm]{Aitchison.pdf}
  \caption{Illustration of the two logratio distance definitions of
    discordance. $d_a$ is the perpendicular Aitchison distance from
    the measured logratio to the concordia line. $d_c$ is the
    Aitchison distance measured along a line connecting the measured
    value and the concordia composition.
  }
  \label{fig:aitchison}
\end{figure}

There exists a vast body of statistical literature detailing the
theoretical and practical advantages of logratio analysis. A deeper
discussion of this topic falls outside the scope of this paper, but
the interested reader is referred to \citet{aitchison1986} and
\citet{pawlowsky2015} for further information. The Euclidean distance
between logratios is also known as the `Aitchison distance'.  We can
redefine discordance as the Aitchison distance from the measured
logratios to the concordia line. We here introduce two ways to do so.
A first option is to simply measure the distance along a perpendicular
line to the concordia curve (Figure~\ref{fig:aitchison}):
\begin{equation}
  d_{a} = dx(t_{68})
  \sin\!\left(\arctan\!\left[ \frac{dy(t_{76})}{dx(t_{68})} \right]
  \right)
  \label{eq:dperp}
\end{equation}

\noindent where
\begin{equation}
\begin{split}
  dx(t)  = & \log\!\left[\frac{{}^{206}Pb}{{}^{238}U}\right]-
  \log\!\left[\exp(\lambda_{238} t)-1\right] \\
  \mbox{~and~}
  dy(t) = & \log\!\left[\frac{{}^{207}Pb}{{}^{206}Pb}\right] -
      \log\!\left[\frac{{}^{235}U}{{}^{238}U}
        \frac{\exp(\lambda_{235}t)-1}{\exp(\lambda_{238}t)-1}
        \right]
  \end{split}
\end{equation}

This definition produces a parallel band around the concordia line in
logarithmic Tera-Wasserburg space. In contrast with $d_r$, $d_t$,
$d_{sk}$, the $d_a$ criterion is less strict at both the young and old
extremes of the geological timescale, and more strict during the
Proterozoic Era, when the U--Pb method is most reliable
(Figure~\ref{fig:dperp}).

The perpendicular Aitchison distance criterion does not take into
account the analytical precision of the isotopic measurements.  To
address this issue, we can also measure the Aitchison distance along a
line connecting the measured logratio and the maximum likelihood
composition on the concordia line:
\begin{equation}
  d_c = \mbox{sgn}[t_{76}-t_{68}] \sqrt{ dx(t_c)^2 + dy(t_c)^2 }
  \label{eq:da}
\end{equation}

\noindent where sgn[$\ast$] stands for the ``sign of $\ast$'', which
produces positive values for measurements that plot above the
concordia line, and negative values for measurements that plot below
it.

\begin{figure}
  \includegraphics[width=8.3cm]{TW-option-4.pdf}
  \caption{The $d_a$ criterion produces a parallel band to the
    (logarithmic) Tera-Wasserburg concordia line. The dashed line is
    plotted at an Aitchison distance of 0.1 from concordia, marking
    a 10\% discordance cutoff. The bottom panel presents the age
    discordance along this line. This shows that the $d_a$ criterion
    is lenient for Phanerozoic and Archaean age estimates, and more
    strict for Proterozoic grains.
  }
  \label{fig:dperp}
\end{figure}

\section{Application to a compilation of detrital zircon U--Pb data}
\label{sec:application}

It is difficult to ascertain the mechanism causing discordance in any
particular zircon grain. Therefore, it is unclear which of the
definitions in Sections~\ref{sec:discordance1} and
\ref{sec:discordance2} is `correct'. All we can is do is apply the
methods to real samples and investigate their outcomes. This Section
will apply the six discordance filters to a compilation of 10,000
detrital zircon U--Pb analyses that were acquired at the London
Geochronology Centre (LGC) at University College London.

The data come from field areas on all seven continents. They were
acquired by LA-ICP-MS, on either an Agilent 7700x (old samples) or an
Agilent 7900 (recent samples) instrument, which was coupled to a New
Wave NWR193 excimer laser. Analytical conditions varied slightly
between different samples but generally used a 25 or 35 micron laser
spot, Ple\v{s}ovice zircon as a primary standard \citep{slama2008}
and, in most cases, GJ-1 zircon as a secondary standard
\citep{jackson2004}. Data reduction was done with GLITTER
\citep{griffin2008}. The data were not subjected to any common Pb
correction or other filters, apart from a visual inspection to remove
the most extreme outliers ($<1\%$ of the data).

Figure~\ref{fig:KDE} shows the frequency distribution of the complete,
unfiltered dataset as a kernel density estimate. The
\textsuperscript{207}Pb/\textsuperscript{235}U,
\textsuperscript{206}Pb/\textsuperscript{238}U and concordia age
spectra all look similar. However, the
\textsuperscript{207}Pb/\textsuperscript{206}Pb age distribution
deviates from the other three chronometers. It reduces the prominence
of the young age components, and inflates the old end of the age
spectrum.

\begin{figure}
  \includegraphics[width=8.3cm]{KDE.pdf}
  \caption{Four superimposed kernel density estimates (with 60~Ma
    bandwidth) for 10,000 unfiltered detrital zircon U--Pb
    dates. The \textsuperscript{207}Pb/\textsuperscript{235}U,
    \textsuperscript{206}Pb/\textsuperscript{238}U and concordia age
    spectra are similar.  However the KDE of the
    \textsuperscript{207}Pb/\textsuperscript{206}Pb data stands
    apart from the other three curves. It deviates both at the young
    end of the age spectrum (which it suppresses), and at the old
    end (which it inflates).
  }
  \label{fig:KDE}
\end{figure}

Figure~\ref{fig:KDEs} applies the six discordance filters to this
database. In order to emphasise the difference between the six
discordance definitions whilst treating them on an equal footing, each
of the filters was adjusted until half of the data were removed. This
was achieved by discordance cutoffs of $d_t$ = 51~Myr, $d_r = 0.0618$,
$d_{p} = 0.084$, $d_{sk} = 0.00225$, $d_a = 0.0219$ and $d_c =
0.0297$.

There are noticeable differences between the density estimates.  As
expected from the theoretical considerations laid out in
Sections~\ref{sec:discordance1} and \ref{sec:discordance2}, the
relative age filter greatly suppresses the younger age components
($<1.5$~Ga) relative to the older parts of the age spectrum
($>1.5$~Ga). The \citet{stacey1975} filter has the opposite effect.
It suppresses the Archaean age component by $\sim$20\% whilst further
increasing the prominence of the Neoproterozoic and Phanerozoic modes.

The discordance definitions based on the absolute age difference and
logratio distances have a comparatively minor effect on the shape of
the age spectrum. All six discordance filters reduce the difference
between the \textsuperscript{207}Pb/\textsuperscript{206}Pb and
concordia age spectra except for the p-value filter, which actually
exacerbates the problem.

If the degree of similarity between the four chronometers is taken as
a measure of their success, then the concordia distance filter ($d_c$)
is the most successful discordance criterion. Apart from a minor
difference at the youngest end of the age spectrum, the four different
chronometers yield virtually identical age distributions after passing
through this filter.

\begin{figure*}[t]
  \includegraphics[width=12cm]{KDEs.pdf}
  \caption{Filtered U--Pb age spectra for the test data, removing the
    50\% most discordant grains according to the six discordance
    filters reviewed in this paper. Notice how the concordia distance
    filter ($d_c$, fifth panel) exhibits the smallest difference
    between the four dating methods (including
    \textsuperscript{207}Pb/\textsuperscript{206}Pb), wherease the
    p-value filter ($d_p$, sixth panel) generates the greatest
    differences between them.  This observation suggests that the
    $d_c$ criterion is the best and the $d_p$ criterion the worst of
    the six methods.
  }
  \label{fig:KDEs}
\end{figure*}

Figure~\ref{fig:KDEs} removed 50\% of the data, in order to emphasise
the differences between the six filters. In real applications, less
stringents discordance filters are usually applied. As mentioned in
the introduction, most current detrital zircon studies apply a
10\%--30\% relative age cutoff.  Using the test data, we can evaluate
the equivalent values for the $d_t$, $d_t$, $d_{sk}$, $d_a$ and $d_c$
criteria (Table~\ref{tab:conversiontable}). For example, a relative
age filter of 10\% removes the same fraction of the test data as an
absolute age filter with $d_t=83.5$~Ma, a Stacey-Kramers filter with
$d_{sk}=0.381$\%, a perpendicular Aitchison filter with $d_a=3.61$\%,
and a concordia distance filter with $d_c=3.67$\%.

The p-value discordance filter has been omitted from this comparison
for two reasons. First, the use of this filter is discouraged for
reasons that have been discussed before. Second, the p-value cutoffs
that are equivalent to any given relative age difference are highly
lab-dependent, with precise equipment requiring different
$d_p$-cutoffs than imprecise instruments. The other five discordance
filters are more universally applicable. So using a different set of
test data should only make a relatively minor difference to the values
in Table~\ref{tab:conversiontable}.

\begin{table}
  \begin{tabular}{c|cccc}
    $d_r$ & $d_t$ & $d_{sk}$ & $d_a$ & $d_c$ \\ \hline
    -10 & -60.4 & -0.287 & -2.59 & -2.62 \\
    -5 & -38.8 & -0.175 & -1.67 & -1.69 \\
    -4 & -32.3 & -0.143 & -1.39 & -1.41 \\
    -3 & -25 & -0.108 & -1.08 & -1.09 \\
    -2 & -17.3 & -0.0681 & -0.747 & -0.758 \\
    -1 & -8.33 & -0.0328 & -0.356 & -0.361 \\
    0 & 0 & 0 & 0 & 0 \\
    1 & 9.31 & 0.0367 & 0.395 & 0.406 \\
    2 & 18.5 & 0.0773 & 0.795 & 0.816 \\
    3 & 28.1 & 0.117 & 1.2 & 1.23 \\
    4 & 36.7 & 0.151 & 1.58 & 1.6 \\
    5 & 43.9 & 0.189 & 1.9 & 1.94 \\
    10 & 83.5 & 0.381 & 3.61 & 3.67 \\
    15 & 117 & 0.538 & 5.02 & 5.16 \\
    20 & 151 & 0.736 & 6.54 & 6.7 \\
    25 & 187 & 0.949 & 8.01 & 8.24 \\
    30 & 230 & 1.18 & 9.93 & 10.2 \\
    40 & 328 & 1.74 & 14.2 & 14.8 \\
    50 & 463 & 2.47 & 20.3 & 21.6
  \end{tabular}
  \caption{Conversion table for the different discordance filters,
    constructed using the test data. All discordance values are
    expressed as \%, except for $d_t$, which is expressed in Ma. This
    table allows the reader to select a discordance cutoff that
    removes the same fraction of their data as the relative age cutoff
    ($d_r$) that they may have applied in the past. }
  \label{tab:conversiontable}
\end{table}

\section{Implementation in \texttt{IsoplotR}}
\label{sec:IsoplotR}

All the discordance filters presented in this paper (both old and new)
have been implemented in \texttt{IsoplotR} \citep{vermeesch2018c}, a
geochronological toolbox written in the \texttt{R}
language. \texttt{IsoplotR} can be accessed either from the command
line, or via a graphical user interface (GUI), either offline or
online (\texttt{http://isoplotr.london-geochron.com}).  The
discordance filters are accessible via both methods. In the GUI, the
discordance can be tabulated via the \texttt{age} function, and has
also been incorporated in \texttt{IsoplotR}'s other functions,
including its concordia and weighted mean calculation algorithms.
Further details about these options are provided under the
\texttt{options} menu. To access the same functionality from the
command line, we must first install \texttt{IsoplotR} from the
Comprehensive R Archive Network (CRAN):

\begin{verbatim}
install.packages('IsoplotR')
\end{verbatim}

\noindent Once installed, we need to add the package to our working
environment:

\begin{verbatim}
library(IsoplotR)
\end{verbatim}

\noindent Loading the test data into memory:

\begin{verbatim}
UPb <- read.data('data.csv',method='U-Pb',format=2)
\end{verbatim}

\noindent Now we can calculate the discordance using
\texttt{IsoplotR}'s \texttt{discfilter} function. For example, to
compute the relative age discordance ($d_r$):

\begin{verbatim}
tr <- age(UPb,discordance=discfilter(option='r'))
\end{verbatim}

\noindent which produces a ${10000}\times{9}$ table whose first eight
columns list the \textsuperscript{207}Pb/\textsuperscript{235}U,
\textsuperscript{206}Pb/\textsuperscript{238}U,
\textsuperscript{207}Pb/\textsuperscript{206}Pb and concordia ages and
their uncertainties, and whose ninth column lists the relative age
discordance as percentages. Similarly, to compute the concordia
distance ($d_c$):

\begin{verbatim}
tc <- age(UPb,discordance=discfilter(option='c'))
\end{verbatim}

\noindent Plotting a KDE of the single grain concordia ages that pass
the perpendicular Aitchison filter with ${-1.67}\leq{d_a}\leq{5.02}$:

\begin{verbatim}
df <- discfilter(option='c',cutoff=c(-1.69,5.16))
kde(UPb,type=5,cutoff.disc=df)
\end{verbatim}

\noindent Apply a Stacey-Kramers common Pb-correction to the data after
applying a Stacey-Kramers discordance filter with
${0.285}\leq{d_{sk}}\leq{1.16}$:

\begin{verbatim}
df <- discfilter(option='sk',cutoff=c(-0.285,1.16))
kde(UPb,common.Pb=3,cutoff.disc=df)
\end{verbatim}

\noindent If the dataset includes \textsuperscript{204}Pb (which is
not the case for the test data), then we can also apply a discordance
filter \textit{after} the common Pb correction. For example:

\begin{verbatim}
df <- discfilter(option='r',before=FALSE,cutoff=c(-5,15))
kde(UPb,common.Pb=3,type=4,cutoff.76=1200,cutoff.disc=df)
\end{verbatim}

\noindent where \verb|option='r'| triggers the relative age filter
($d_r$), \texttt{common.Pb=3} applies a Stacey-Kramers type common Pb
correction, \texttt{type=4} uses the
\textsuperscript{206}Pb/\textsuperscript{238}U-age for young grains
and the \textsuperscript{207}Pb/\textsuperscript{206}Pb-age for old
ones, and \texttt{cutoff.76} marks the age (in Ma) at which to switch
from the \textsuperscript{206}Pb/\textsuperscript{238}U to the
\textsuperscript{207}Pb/\textsuperscript{206}Pb method. Further
information about these functions can be obtained from the built-in
documentation:

\begin{verbatim}
?IsoplotR
?discfilter
?kde
\end{verbatim}

\noindent Note that the examples shown here may take a few minutes to
complete due to the large size of the test dataset.

\conclusions
\label{sec:conclusions}

This paper compared four U--Pb clocks and six discordance filters.

\begin{enumerate}
  \item The \textsuperscript{206}Pb/\textsuperscript{238}U clock is
    most precise at the young end of the geologic timescale.
  \item The \textsuperscript{207}Pb/\textsuperscript{206}Pb method is
    more precise than the
    \textsuperscript{206}Pb/\textsuperscript{238}U method before the
    Neoproterozoic.
  \item The \textsuperscript{207}Pb/\textsuperscript{235}U clock
    offers no advantage over the other two methods.
  \item The single grain concordia age is applicable to the entire
    span of geologic time and always offers the best precision. It
    approaches the \textsuperscript{206}Pb/\textsuperscript{238}U age
    as time approaches zero, and the
    \textsuperscript{207}Pb/\textsuperscript{206}Pb age as time
    approaches infinity.
\end{enumerate}

\noindent The six discordance filters include three existing ones and
three new ones.

\begin{enumerate}
  \item The relative age discordance $d_r$ is the most widely used
    criterion today. It is more likely to remove young grains than old
    ones.
  \item The absolute age discordance $d_t$ is not widely used. But it
    illustrates the dramatic effect that the discordance definition
    can have on the filtered age distibutions. In contrast with the
    relative age filter, it is most likely to reject old grains, and
    least likely to reject young ones.
  \item The p-value based discordance filter $d_p$ may have intuitive
    appeal as an `objective' definition. But it has an undesirable
    negative effect on the precision and accuracy of the filtered
    results.
  \item The Stacey-Kramers discordance filter $d_{sk}$ assumes that
    discordance is solely caused by common Pb contamination. If this
    assumption is correct, then the $d_{sk}$ filter will produce the
    most accurate age distributions, provided that a
    \citet{stacey1975} common Pb correction is applied to the filtered
    data.
  \item The perpendicular Aitchison distance $d_a$ is a useful vehicle
    to illustrate the application of logratio statistics to detrital
    zircon U--Pb geochronology. It produces a parallel acceptance zone
    around the (log-transformed) concordia line. This filter is most
    likely to reject `middle aged' zircon grains, between 1000 and
    2000~Ma, where the age resolving power of the U--Pb method is
    greatest. Above and below this interval, the $d_a$ criterion is
    more forgiving. This behaviour is desirable because natural
    samples tend to exhibit more age discordance below 1000~Ma and
    above 2000~Ma than between these dates.
  \item The concordia distance $d_c$ is a modified version of the
    $d_a$ criterion that takes into account the (correlated)
    uncertainties of the U--Pb isotopic composition.  Its effects on
    the U--Pb age distributions are more difficult to visualise but
    are similar to those of the $d_a$ criterion.  Applying the $d_c$
    filter to the test data shows that it minimises the difference
    between the \textsuperscript{207}Pb/\textsuperscript{235}U,
    \textsuperscript{206}Pb/\textsuperscript{238}U,
    \textsuperscript{207}Pb/\textsuperscript{206}Pb and concordia age
    spectra. Therefore the $d_c$ criterion simplifies the
    interpretation of detrital zircon U--Pb age spectra.
\end{enumerate}

%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.

%\codeavailability{TEXT} %% use this section when having only software code available

%\dataavailability{TEXT} %% use this section when having only data sets available

\codedataavailability{\texttt{IsoplotR} is free software released
  under the GPL-3 license. The package and its source code are
  available from
  \texttt{https://cran.r-project.org/package=IsoplotR}. The test data
  can be downloaded from
  \texttt{https://github.com/pvermees/discpaper/}.}

%\sampleavailability{TEXT} %% use this section when having geoscientific samples available

%\videosupplement{TEXT} %% use this section when having video supplements available

\appendix

\section{Comparing the precision of the
  \textsuperscript{207}Pb/\textsuperscript{235}U,
  \textsuperscript{206}Pb/\textsuperscript{238}U,
  \textsuperscript{207}Pb/\textsuperscript{206}Pb and concordia
  clocks}\label{app}

The uncertainty of a U--Pb date depends on three factors:

\begin{enumerate}
\item the age and, hence, the true isotopic ratio;
\item the sensitivity of the ion detectors to U and Pb; and
\item the dwell times used to measure the different isotopes.
\end{enumerate}

These three factors vary between samples, and between labs.  In order
to explore their effects, let us first define the following
parameters:

\begin{itemize}
\item $t_{68}$, $t_{75}$ and $t_{76}$: the
  \textsuperscript{206}Pb/\textsuperscript{238}U,
  \textsuperscript{207}Pb/\textsuperscript{235}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb ages (in Ma);
\item $\lambda_{38}$ and $\lambda_{35}$: the decay constants of
  \textsuperscript{238}U and \textsuperscript{235}U (in
  Ma\textsuperscript{-1});
\item $R_{85}$: the natural
  \textsuperscript{238}U/\textsuperscript{235}U ratio;
\item $R_{68}$, $R_{75}$ and $R_{76}$: the true
  \textsuperscript{206}Pb/\textsuperscript{238}U,
  \textsuperscript{207}Pb/\textsuperscript{235}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb atomic ratios;
\item $r_{68}$, $r_{75}$ and $r_{76}$: the measured
  \textsuperscript{206}Pb/\textsuperscript{238}U,
  \textsuperscript{207}Pb/\textsuperscript{235}U and
  \textsuperscript{207}Pb/\textsuperscript{206}Pb signal ratios;
\item $f^{Pb}_{U}$: the fractionation factor between Pb and U;
\item $d^{06}_{38}$: the dwell time ratio of \textsuperscript{206}Pb
  and \textsuperscript{238}U;
\item $d^{07}_{06}$: the dwell time ratio of \textsuperscript{207}Pb
  and \textsuperscript{206}Pb;
\item $n_{06}$, $n_{07}$ and $n_{38}$: the number of
  \textsuperscript{206}Pb, \textsuperscript{207}Pb and
  \textsuperscript{238}U ions counted during a measurement.
\end{itemize}

Then the true isotope ratios are given by:

\begin{equation}
  R_{68} = \exp(\lambda_{38}t_{68})-1
\end{equation}
\begin{equation}
  R_{75} = \exp(\lambda_{35}t_{75})-1
\end{equation}
\begin{equation}  
   R_{76} = \frac{1}{R_{85}}\frac{R_{75}}{R_{68}}
\end{equation}

and the measured ratios by:

\begin{equation}
    r_{68} = d^{06}_{38}f^{Pb}_{U}R_{68}
\end{equation}
\begin{equation}
    r_{75} = d^{07}_{06}d^{06}_{38}f^{Pb}_{U}R_{75}
\end{equation}
\begin{equation}  
    r_{76} = d^{07}_{06}R_{76}
\end{equation}

so that the predicted \textsuperscript{206}Pb and
\textsuperscript{207}Pb ion counts can be written as:
\begin{equation}
    n_{06} = n_{38}d^{06}_{38}f^{Pb}_{U}R_{68}
\end{equation}
\begin{equation}  
    n_{07} = n_{06}d^{07}_{06}R_{76}
\end{equation}

Assuming that all the ions are measured by Secondary Electron
Multiplier (SEM), with analytical uncertainties that are governed by
Poissonian shot noise:
\begin{equation}
    \left(\frac{\sigma[r_{68}]}{r_{68}}\right)^2 =
    \frac{1}{n_{38}} + \frac{1}{n_{06}}
\end{equation}
\begin{equation}  
    \left(\frac{\sigma[r_{75}]}{r_{75}}\right)^2 =
    \frac{1}{n_{38}} + \frac{1}{n_{07}}
\end{equation}
\begin{equation}  
    \left(\frac{\sigma[r_{76}]}{r_{76}}\right)^2 =
    \frac{1}{n_{06}} + \frac{1}{n_{07}}
\end{equation}

then the standard errors of the signal ratios ratios are given by:
\begin{equation}
    \sigma[r_{68}] = \frac{n_{06}}{n_{38}}
    \sqrt{\frac{1}{n_{38}} + \frac{1}{n_{06}}}
\end{equation}
\begin{equation}  
    \sigma[r_{75}] = R_{85} \frac{n_{07}}{n_{38}}
    \sqrt{\frac{1}{n_{38}} + \frac{1}{n_{07}}}
\end{equation}
\begin{equation}  
    \sigma[r_{76}] = \frac{n_{07}}{n_{06}}
    \sqrt{\frac{1}{n_{06}} + \frac{1}{n_{07}}}
\end{equation}

Finally, the uncertainties of the age estimates are given by standard
error propagation:
\begin{equation}
    \sigma[t_{68}] = \frac{\partial{t_{68}}}{\partial{r_{68}}} \sigma[r_{68}]
\end{equation}
\begin{equation}  
    \sigma[t_{75}] = \frac{\partial{t_{75}}}{\partial{r_{75}}} \sigma[r_{75}]
\end{equation}
\begin{equation}  
    \sigma[t_{76}] = \frac{\partial{t_{76}}}{\partial{r_{76}}} \sigma[r_{76}]
\end{equation}

where
\begin{equation}
    \frac{\partial{t_{68}}}{\partial{r_{68}}} = 
    \frac{1}{\lambda_{38}(1+R_{68})} \frac{1}{d^{06}_{38}f^{Pb}_{U}}
\end{equation}
\begin{equation}  
    \frac{\partial{t_{75}}}{\partial{r_{75}}} = 
    \frac{1}{\lambda_{35}(1+R_{75})} \frac{1}{d^{07}_{06}f^{Pb}_{U}}
\end{equation}
\begin{equation}  
  \frac{\partial{t_{76}}}{\partial{r_{76}}} =
  \frac{R_{85}R_{68}^2}
       {(\partial{R_{75}}/\partial{t_{75}})R_{68} -
         R_{75}(\partial{R_{68}}/\partial{t_{68}})}
       \frac{1}{d^{07}_{06}}
\end{equation}

Figure~\ref{fig:precision} shows the result of these calculations
using realistic values of $n_{38}$, $f^{Pb}_{U}$ and $d^{07}_{06}$,
which yield an outcome that is similar to the test data.

\noappendix       %% use this to mark the end of the appendix section. Otherwise the figures might be numbered incorrectly (e.g. 10 instead of 1).

%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:

%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
%% They will be correctly named automatically.

%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:

%\appendixfigures  %% needs to be added in front of appendix figures

%\appendixtables   %% needs to be added in front of appendix tables

%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.

\authorcontribution{This is a single author publication.} %% this section is mandatory

\competinginterests{The author has no competing interests.} %% this section is mandatory even if you declare that no competing interests are present

%\disclaimer{TEXT} %% optional section

\begin{acknowledgements}
  The writing of this paper was triggered by a stimulating email
  conversation with Chris Spencer and Steve Puetz. This research was
  supported by NERC standard grant \#NE/T001518/1 (`Beyond Isoplot').
\end{acknowledgements}

%% REFERENCES

%% The reference list is compiled as follows:

\bibliographystyle{copernicus}
\bibliography{/home/pvermees/Dropbox/biblio.bib}

%\begin{thebibliography}{}

%\bibitem[AUTHOR(YEAR)]{LABEL1}
%REFERENCE 1

%\bibitem[AUTHOR(YEAR)]{LABEL2}
%REFERENCE 2

%\end{thebibliography}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
%% \bibliographystyle{copernicus}
%% \bibliography{example.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

%% FIGURES

%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
%% between bibliography and first table and/or figure as well as between each table and/or figure.

% The figure files should be labelled correctly with Arabic numerals (e.g. fig01.jpg, fig02.png).

%% ONE-COLUMN FIGURES

%%f
%\begin{figure}[t]
%\includegraphics[width=8.3cm]{FILE NAME}
%\caption{TEXT}
%\end{figure}
%
%%% TWO-COLUMN FIGURES
%
%%f
%\begin{figure*}[t]
%\includegraphics[width=12cm]{FILE NAME}
%\caption{TEXT}
%\end{figure*}
%
%
%%% TABLES
%%%
%%% The different columns must be seperated with a & command and should
%%% end with \\ to identify the column brake.
%
%%% ONE-COLUMN TABLE
%
%%t
%\begin{table}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table}
%
%%% TWO-COLUMN TABLE
%
%%t
%\begin{table*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table*}
%
%%% LANDSCAPE TABLE
%
%%t
%\begin{sidewaystable*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{sidewaystable*}
%
%
%%% MATHEMATICAL EXPRESSIONS
%
%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
%%%
%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
%%% Vectors are identified in bold italic font using \vec{x}
%%% Matrices are identified in bold roman font
%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
%%% The character * should not be applied as mutliplication sign
%
%
%%% EQUATIONS
%
%%% Single-row equation
%
%\begin{equation}
%
%\end{equation}
%
%%% Multiline equation
%
%\begin{align}
%& 3 + 5 = 8\\
%& 3 + 5 = 8\\
%& 3 + 5 = 8
%\end{align}
%
%
%%% MATRICES
%
%\begin{matrix}
%x & y & z\\
%x & y & z\\
%x & y & z\\
%\end{matrix}
%
%
%%% ALGORITHM
%
%\begin{algorithm}
%\caption{...}
%\label{a1}
%\begin{algorithmic}
%...
%\end{algorithmic}
%\end{algorithm}
%
%
%%% CHEMICAL FORMULAS AND REACTIONS
%
%%% For formulas embedded in the text, please use \chem{}
%
%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
%
%\begin{reaction}
%%% \rightarrow should be used for normal (one-way) chemical reactions
%%% \rightleftharpoons should be used for equilibria
%%% \leftrightarrow should be used for resonance structures
%\end{reaction}
%
%
%%% PHYSICAL UNITS
%%%
%%% Please use \unit{} and apply the exponential notation

\end{document}